MSstatsPTM: Statistical relative quantification of post-translational modifications in bottom-up mass spectrometry-based proteomics

### 2022-07-30 main/MSstatsPTM_Main.tex
Small: discussion

\todo{When you have a chance, could you do minor clean ups of the references? Remove the doi; check that the same journal has the same abbreviation in all the references; keep the reference format consistent (e.g., remove capitalization in article titles; italic for book/volume titles etc)} We proposed a general statistical modeling framework and implementation for PTM characterization. The framework is designed for bottom-up MS workflows, which are characterized with variations from multiple confounded sources, frequent missing data, and associated uncertainty in the conclusions. The framework is general and is applicable to a variety of experimental designs. It outperforms the ad-hoc methods underlying ANOVA and Limma, and yields accurate results in the broad type of experimental circumstances, including the presence of missing values, changes in protein abundance, few representative peptides, and different acquisition methods. The framework allows us to plan for subsequent experiments, and choose the appropriate number of replicates in consideration of adjustment with respect to protein abundance. The implementation allows for straightforward application of the methods discussed and allows for reproducible experimental analysis.

>>>>

\todo{When you have a chance, could you do minor clean ups of the references? Remove the doi; check that the same journal has the same abbreviation in all the references; keep the reference format consistent (e.g., remove capitalization in article titles; italic for book/volume titles etc)}

\todo{Somewhere in the discussion it would be good to comment that the proposed approach relies on the commutative property of the null hypothesis in Eq 2. While the ratio-based approaches look at the differences within one sample first, we look at the differences between conditions first. This provides us with a greater flexibility as far as modeling, accounting for outliers and missing values, and handling complex designs and robust implementations.}

We proposed a general statistical modeling framework and implementation for PTM characterization. The framework is designed for bottom-up MS workflows, which are characterized with variations from multiple confounded sources, frequent missing data, and associated uncertainty in the conclusions. The framework is general and is applicable to a variety of experimental designs. It outperforms the ad-hoc methods underlying ANOVA and Limma, and yields accurate results in the broad type of experimental circumstances, including the presence of missing values, changes in protein abundance, few representative peptides, and different acquisition methods. The framework allows us to plan for subsequent experiments, and choose the appropriate number of replicates in consideration of adjustment with respect to protein abundance. The implementation allows for straightforward application of the methods discussed and allows for reproducible experimental analysis.
